Macro Impacts

### 2016-08-24 Draft.tex

\section{Seismic Wave Detection}
The seismometers on the Moon are sensitive to displacements as small as $0.3\te{\,nm}$. We calculate the displacement due to each ray, and describe how they add. Consider a $p$-wave traveling towards positive $x$ and of compact support $S$ in the plane normal to to its motion. Denote $A =\int_S\diff y\diff z$ The displacement field is
\begin{align}
u(\vec x,t) &= \chi_{\R\times S}\int\f{\diff k}{2\pi}U(k)e^{-\I k(x-v_p t)}\,,\\
\abs{u(\vec x,t)}&\leq\int\f{\diff k}{2\pi}\abs{U(k)}\,,
\end{align}
and the energy is
\begin{align}
E = \rho v_p^2 A\int\f{\diff k}{2\pi}k^2\abs{U(k)}^2\,.
\end{align}
As before, we denote
\begin{align}
E_k = \rho v_p^2 A\int_0^k\f{\diff k'}{2\pi}k'^2\abs{U(k')}^2\,.
\end{align}
It follows that
\begin{align}
\abs{U(k)} = \sqrt{\f{2\pi}{\rho v_p^2 A}}\f1k\sqrt{\td{E_k}{k}{}}\,,
\end{align}
and from the estimate above
\begin{align}
\abs{u(\vec x,t)}\leq(2\pi\rho v_p^2 A)^{-1/2}\int_0^k\f{\diff k'}{k'}\sqrt{\td{E_{k'}}{k'}{}}\,.
\end{align}
In our case $\Xi E= C_1 k^2$ before attenuation. After anelastic attenuation, $E_k =\te{Ref}\,2C_1 k e^{-k\,\te{Att}}$
\begin{align}
\abs{u(\vec x,t)}\leq\sqrt{\f{2}{\rho v_p^2 A}}\sqrt{\f{\te{Ref}\,\Xi E}{\te{Att}}}\f1k\te{Erf}\ps{\sqrt{\f{\te{Att}\,k}{2}}}\,.
\end{align}
\section{Simulation}
The simulation is broken down into two natural parts: data generation and data analysis. The data generation consists of propagating a large number of rays to approximate a line source of $p$-waves and tracking the positions and energies at which they contact the surface of the moon. The data analysis makes use of HEALPix to approximate the energy density on the sphere and then makes use of section IV to measure displacements.
\subsection{Data Generation}
To generate the data, we consider a Macro trajectory which passes a distance $D$ from the center of the moon. This trajectory has a length $L$. Along this trajectory, we consider a set of $M$ points, from each of which we shoot $N$ rays, each endowed with $E_m = \rho(r_{m})\sigma_Xv_X^2 L/(MN)$ energy, where $m\in\{1,\dots,M\}$ labels the point. Because of the spherical symmetry, each ray is confined to a subspace of two dimensions. We use the trajectories in section III to propagate the rays. Since the moon is stratified, we must halt the propagation at the boundary of each layer and reflect/transmit rays. During a ray's propagation through a given stratum $\Delta t_i$ is recorded and added to $\Delta t$, $\te{ATT}_i$ is recorded and added to $\te{ATT}$, and $\te{REF}_i$ is recorded at the boundary and is multiplied by $\te{REF}$. Each time a ray makes contact with the surface of the Moon, its position, $\te{ATT}$, $\te{REF}$, $\Delta t$, and $E_m$ are recorded. The final data is a table of these values corresponding to each ray intersection with the surface. We iterate 16 times since it takes 16 iterations for a ray to pass entirely through the VPREMOON model stratification.
\subsection{Data Analysis}
We now convert the position data to HEALPix coordinates for a given $n_\te{side}$. Define $A_\te{pix} = 4\pi R_\te{moon}^2/12\times 2^{n_{\te{side}}}$, the area of a HEALPix pixel. The time it takes a ray to cross a HEALPix pixel can be estimated $t_\te{crossing}^2 = A_\te{pix}/\pi v_p^2$. Rays that contact a pixel within $t_\te{crossing}$ of one another are considered to add constructively. Since the integral in section IV is unreasonable to do for more than one ray, we make use of Minkowski's inequality and obtain an upper bound on displacement by simply adding the displacements from each ray. We suppose that each ray distributes its energy evenly over the area of the HEALPix pixel, and thus set $A = A_\te{pix}$. This is a good approximation when $N$ is the same as the number of HEALPix pixels.

@@

If the energy of a ray from every point along the line hitting a single pixel (with no reflection, refraction, or attenuation) is less than the displacement threshold, there is no chance of detection and no need to run the simulation. (If two rays from the same point on the line source hit the same pixel as $n_\te{side}\to\infty$, then they must be the same ray).
\begin{align}
u\leq\sqrt{\f{\Xi E}{\pi^2\rho k v_p^2 R_\te{Moon}^2}}
\end{align}
This is independent of the choice of partition. For the triangle wave, we need not test $\sigma_X\leq10^{-9}\te{m}^2$ (I WILL VERIFY THIS).
\section{Results}

>>>>

\section{Seismic Wave Detection}
The seismometers on the Moon are sensitive to displacements as small as $0.3\te{\,nm}$. We calculate the displacement due to each ray, and describe how they add. Consider a $p$-wave traveling towards positive $x$ and of compact support $S$ in the plane normal to to its motion. Denote $A =\int_S\diff y\diff z$ The displacement field is
\begin{align}
u(\vec x,t) &= \chi_{\R\times S}\int\f{\diff k}{2\pi}U(k)e^{-\I k(x-v_p t)}\,,\\
\abs{u(\vec x,t)}&\leq\int\f{\diff k}{2\pi}\abs{U(k)}\,,
\end{align}
and the energy is
\begin{align}
E = \rho v_p^2 A\int\f{\diff k}{2\pi}k^2\abs{U(k)}^2\,.
\end{align}
As before, we denote
\begin{align}
E_k = \rho v_p^2 A\int_0^k\f{\diff k'}{2\pi}k'^2\abs{U(k')}^2\,.
\end{align}
It follows that
\begin{align}
\abs{U(k)} = \sqrt{\f{2\pi}{\rho v_p^2 A}}\f1k\sqrt{\td{E_k}{k}{}}\,,
\end{align}
and from the estimate above
\begin{align}
\abs{u(\vec x,t)}\leq(2\pi\rho v_p^2 A)^{-1/2}\int_0^k\f{\diff k'}{k'}\sqrt{\td{E_{k'}}{k'}{}}\,.
\end{align}
In our case $\Xi E= C_1 k^2$ before attenuation. After anelastic attenuation, $E_k =\te{Ref}\,2C_1 k e^{-k\,\te{Att}}$
\begin{align}
\abs{u(\vec x,t)}\leq\sqrt{\f{2}{\rho v_p^2 A}}\sqrt{\f{\te{Ref}\,\Xi E}{\te{Att}}}\f1k\te{Erf}\ps{\sqrt{\f{\te{Att}\,k}{2}}}\,.
\end{align}
(28) is impossible to integrate when $E_k$ consists of the energies of two different rays, each with different attenuation factors. When this case arrises, we average $\te{Att}$ among the coincident rays. For the lowest frequency modes $f<20\te{Hz}$, $\te{Att} k$ order 10 at most, and does not contribute significantly to $u$. Taking $\te{Att} = 0$, we have
\begin{align}
\abs{u(\vec x,t)}\leq\sqrt{\f{2}{\rho v_p^2 A}}\sqrt{\f{\te{Ref}\,\Xi E}{\te{Att}}}\f1k\te{Erf}\ps{\sqrt{\f{\te{Att}\,k}{2}}}\,.
\end{align}
\section{Simulation}
The simulation is broken down into two natural parts: data generation and data analysis. The data generation consists of propagating a large number of rays to approximate a line source of $p$-waves and tracking the positions and energies at which they contact the surface of the moon. The data analysis makes use of HEALPix to approximate the energy density on the sphere and then makes use of section IV to measure displacements.
\subsection{Data Generation}
To generate the data, we consider a Macro trajectory which passes a distance $D$ from the center of the moon. This trajectory has a length $L$. Along this trajectory, we consider a set of $M$ points, from each of which we shoot $N$ rays, each endowed with $E_m = \rho(r_{m})\sigma_Xv_X^2 L/(MN)$ energy, where $m\in\{1,\dots,M\}$ labels the point. Because of the spherical symmetry, each ray is confined to a subspace of two dimensions. We use the trajectories in section III to propagate the rays. Since the moon is stratified, we must halt the propagation at the boundary of each layer and reflect/transmit rays. During a ray's propagation through a given stratum $\Delta t_i$ is recorded and added to $\Delta t$, $\te{ATT}_i$ is recorded and added to $\te{ATT}$, and $\te{REF}_i$ is recorded at the boundary and is multiplied by $\te{REF}$. Each time a ray makes contact with the surface of the Moon, its position, $\te{ATT}$, $\te{REF}$, $\Delta t$, and $E_m$ are recorded. The final data is a table of these values corresponding to each ray intersection with the surface. We iterate 16 times since it takes 16 iterations for a ray to pass entirely through the VPREMOON model stratification.
\subsection{Data Analysis}
We now convert the position data to HEALPix coordinates for a given $n_\te{side}$. Define $A_\te{pix} = 4\pi R_\te{moon}^2/12\times 2^{n_{\te{side}}}$, the area of a HEALPix pixel. The time it takes a ray to cross a HEALPix pixel can be estimated $t_\te{crossing}^2 = A_\te{pix}/\pi v_p^2$. Rays that contact a pixel within $t_\te{crossing}$ of one another are considered to add constructively. Since the integral in section IV is unreasonable to do for more than one ray, we make use of Minkowski's inequality and obtain an upper bound on displacement by simply adding the displacements from each ray. We suppose that each ray distributes its energy evenly over the area of the HEALPix pixel, and thus set $A = A_\te{pix}$. This is a good approximation when $N$ is the same as the number of HEALPix pixels.

@@

If the energy of a ray from every point along the line hitting a single pixel (with no reflection, refraction, or attenuation) is less than the displacement threshold, there is no chance of detection and no need to run the simulation. (If two rays from the same point on the line source hit the same pixel as $n_\te{side}\to\infty$, then they must be the same ray).
\begin{align}
u\leq\sqrt{\f{\Xi E}{\pi^2\rho k v_p^2 R_\te{Moon}^2}}
\end{align}
This is independent of the choice of partition. For the triangle wave, we need not test $\sigma_X\leq 4\times 10^{-11}\te{m}^2$.
\section{Results}

### 2016-08-24 Macro Impacts Revised.tex

Consider the initial condition of a delta source along the z-axis in cylindrical coordinates:
\begin{align}
T(r,0) = \f{\abs{\frac{dE}{dx}}r_X}{\rho \sigma_X c_p}\delta(r),
\end{align}
Note that this model/initial condition is reasonable given that the delta source becomes a Gaussian with the same energy (assuming energy is conserved) an instant after the initial input of the delta source.

Taking the Fourier transform of this initial condition over r and $\phi$ fixes $C_1$ as 
\begin{align}
C_1 =  2\pi \f{\abs{\frac{dE}{dx}}r_X}{\rho \sigma_X c_p}\int_{0}^{\infty}\delta(r)e^{i k r} dr = 2\pi\f{\abs{\frac{dE}{dx}}r_X}{\rho \sigma_X c_p}\,,
\end{align}
where the factor of $2 \pi$ comes from the $\phi$ integral.
$T$ in real space is given by inverting the Fourier transform:
\begin{align}
T(r,t) &=\frac{1}{\p{2\pi}} \int_0^\infty k_r \diff{k_r}\tilde{T} = \frac{2\pi}{2\pi}\f{\abs{\frac{dE}{dx}}r_X}{\rho \sigma_X c_p}\int_0^\infty k_r\diff k_r e^{-\alpha k_r^2 t}J_0(k_r r)\,,\\
&=\f{\abs{\frac{dE}{dx}}r_X}{2 \rho \sigma_X c_p \alpha t}\ e^{-\f{r^2}{4t\alpha}}\,.
\end{align}
Setting $T = T_\te{melt}$, we have
\begin{align}
r_\te{melt} = \sqrt{4t\alpha\ln\ps{\f{\abs{\frac{dE}{dx}}r_X}{2 \rho \sigma_X c_p \alpha T_\te{melt} t}}}\,.
\end{align}
where we've taken the positive branch of the square root, since we only care about the melt radius increasing. Then taking the derivative w.r.t to time gives
\begin{align}
\dot r_\te{melt} &=\sqrt{\f{\alpha}{t}}\f{\ln\p{\f{\abs{\frac{dE}{dx}}r_X}{2 \rho \sigma_X c_p \alpha T_\te{melt} t}} - 1}{\sqrt{\ln\p{\f{\abs{\frac{dE}{dx}} r_X}{2 \rho \sigma_X c_p \alpha T_\te{melt} t}}}}\,.
\end{align}
Since, for the parameters at hand, $\ln\p{\f{\abs{\frac{dE}{dx}}r_X}{2 \rho \sigma_X c_p \alpha T_\te{melt} t}}$ is at least of order 10 $>>$ 1, it is a reasonable approximation to take
\begin{align}
\dot r_\te{melt} &=\sqrt{\f{\alpha}{t}\ln\p{\frac{\abs{\frac{dE}{dx}}r_X}{2 \rho \sigma_X c_p \alpha T_\te{melt} t}}}\,.
\end{align}
Thus, the time at which the velocity of the melt front is that of the speed of sound, $v_p$, is given by using the Lambert W-function (or "ProductLog" in Mathamatica) as
\begin{align}
t_{\te{fast}} = \f{\alpha}{v_p^2}W\p{{\f{\abs{\frac{dE}{dx}}r_Xv_p^2}{2 \rho \sigma_X c_p \alpha^2 T_\te{melt}}}}
\end{align}
Plugging in $t_{\te{fast}}$ to $r_\te{melt}$, which is a solid approximation for the given case, gives
\begin{align}
r_\te{fast melt} =  \frac{2\alpha}{v_p}\sqrt{W\p{{\f{\abs{\frac{dE}{dx}}r_{X}v_p^2}{2 \rho \sigma_X c_p \alpha^2 T_\te{melt}}}}\ln\ps{\p{\frac{\abs{\frac{dE}{dx}}r_{X}v_p^2}{2 \rho \sigma_X c_p \alpha^2 T_\te{melt}}}\frac{1}{W\p{{\f{\abs{\frac{dE}{dx}}r_{X}v_p^2}{2 \rho \sigma_X c_p \alpha^2 T_\te{melt}}}}
}}}\,.
\end{align}
Notice that this has the form of $\sqrt{W(x)\ln\p{{\frac{x}{W(x)}}}}$, which gives $\sqrt{W(x)^2}$ by the identity $\ln\p{{\frac{x}{W(x)}}} = W(x)$ which holds when $x \geq -1/e$ which is true for the parameters at hand.  Thus, we have,
\begin{align}
r_\te{fast melt} =\frac{2\alpha}{v_p}W\p{{\f{\abs{\frac{dE}{dx}}r_{X}v_p^2}{2 \rho \sigma_X c_p \alpha^2 T_\te{melt}}}} = \frac{2\alpha}{v_p}W\p{\frac{r_{X}v_{X} v_p^2}{2 c_p \alpha^2 T_\te{melt}}}.
\end{align}
which is on the order of micrometers to tenths of micrometers for most marco radii which, for $r_X > 10^-8$ results in $r_\te{fast melt} < r_X$ which is something of a problem.

This either suggests that much of the melting occurs once the temperature field is expanding slower than the speed of sound, and that the expansion speed of the melt zone slows down very quickly or that the material parameters -- especially $\alpha$ -- are far from constant (or may not even be well defined) in our case.

>>>>

Consider the initial condition of a delta source along the z-axis in cylindrical coordinates:
\begin{align}
T(r,0) = \f{\abs{\frac{dE}{dx}}}{\rho \sigma_x c_p}\delta(r),
\end{align}
Taking the Fourier transform of this initial condition over r and $\phi$ fixes $C_1$ as 
\begin{align}
C_1 =  2\pi \f{\abs{\frac{dE}{dx}}}{\rho \sigma_x c_p}\int_{0}^{\infty}\delta(r)e^{i k r} dr = 2\pi\f{\abs{\frac{dE}{dx}}}{\rho \sigma_x c_p}\,,
\end{align}
where the factor of $2 \pi$ comes from the $\phi$ integral.
$T$ in real space is given by inverting the Fourier transform:
\begin{align}
T(r,t) &=\frac{1}{\p{2\pi}} \int_0^\infty k_r \diff{k_r}\tilde{T} = \frac{2\pi}{2\pi}\f{\abs{\frac{dE}{dx}}}{\rho \sigma_x c_p}\int_0^\infty k_r\diff k_r e^{-\alpha k_r^2 t}J_0(k_r r)\,,\\
&=\f{\abs{\frac{dE}{dx}}}{2 \rho \sigma_x c_p \alpha t}\ e^{-\f{r^2}{4t\alpha}}\,.
\end{align}
Setting $T = T_\te{melt}$, we have
\begin{align}
r_\te{melt} = \sqrt{4t\alpha\ln\ps{\f{\abs{\frac{dE}{dx}}}{2 \rho \sigma_x c_p \alpha T_\te{melt} t}}}\,.
\end{align}
where we've taken the positive branch of the square root, since we only care about the melt radius increasing. Then taking the derivative w.r.t to time gives
\begin{align}
\dot r_\te{melt} &=\sqrt{\f{\alpha}{t}}\f{\ln\p{\f{\abs{\frac{dE}{dx}}}{2 \rho \sigma_x c_p \alpha T_\te{melt} t}} - 1}{\sqrt{\ln\p{\f{\abs{\frac{dE}{dx}}}{2 \rho \sigma_x c_p \alpha T_\te{melt} t}}}}\,.
\end{align}
Since, for the parameters at hand, $\ln\p{\f{\abs{\frac{dE}{dx}}}{2 \rho \sigma_x c_p \alpha T_\te{melt} t}}$ is at least of order 10 $>>$ 1, it is a reasonable approximation to take
\begin{align}
\dot r_\te{melt} &=\sqrt{\f{\alpha}{t}\ln\p{\f{\abs{\frac{dE}{dx}}}{2 \rho \sigma_x c_p \alpha T_\te{melt} t}}}\,.
\end{align}
Thus, the time at which the velocity of the melt front is that of the speed of sound, $v_p$, is given by using the Lambert W-function (or "ProductLog" in Mathamatica) as
\begin{align}
t_{\te{fast}} = \f{\alpha}{v_p^2}W\p{{\f{\abs{\frac{dE}{dx}}v_p^2}{2 \rho \sigma_x c_p \alpha^2 T_\te{melt}}}}
\end{align}
Plugging in $t_{\te{fast}}$ to $r_\te{melt}$, which is a solid approximation for the given case, gives
\begin{align}
r_\te{fast melt} =  \frac{2\alpha}{v_p}\sqrt{W\p{{\f{\abs{\frac{dE}{dx}}v_p^2}{2 \rho \sigma_x c_p \alpha^2 T_\te{melt}}}}\ln\ps{\p{\frac{\abs{\frac{dE}{dx}}v_p^2}{2 \rho \sigma_x c_p \alpha^2 T_\te{melt}}}\frac{1}{W\p{{\f{\abs{\frac{dE}{dx}}v_p^2}{2 \rho \sigma_x c_p \alpha^2 T_\te{melt}}}}
}}}\,.
\end{align}
Notice that this has the form of $\sqrt{W(x)\ln\p{{\frac{x}{W(x)}}}}$, which gives $\sqrt{W(x)^2}$ by the identity $\ln\p{{\frac{x}{W(x)}}} = W(x)$ which holds when $x \geq -1/e$ which is true for the parameters at hand.  Thus, we have,
\begin{align}
r_\te{fast melt} =\frac{2\alpha}{v_p}W\p{{\f{r_X\abs{\frac{dE}{dx}}v_p^2}{2 \rho \sigma_x c_p \alpha^2 T_\te{melt}}}} = \frac{2\alpha}{v_p}W\p{\frac{r_Xv_{X} v_p^2}{2 c_p \alpha^2 T_\te{melt}}}.
\end{align}
which is on the order of micrometers to tenths of micrometers for our case.

Conceptually, it would seem that $r_\te{fast melt}$ should be added to $r_\te{Macro}$ to give the full extent of material melted down.
